First paper - Draft

### 2017-06-16 paper/paper.Rnw
Moving code into paper - 30% done

<<packages_conf, echo = F>>=
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # For pisa_countrynames
  library(countrycode) # For region variable
  library(car)
  library(readr)
  library(SAScii)
  library(inequalityintsvy)
  library(lme4)
  library(modelr)
  library(tidyverse)

@@

@@

<<loading_data-recoding>>=
pisa_all <- read_rds("./data/pisa_listcol.Rdata")
pisa_all2 <- pisa_all

years <- seq(2000, 2015, 3)
countries <- c("Germany", "United States", "Denmark", "Sweden", "United Kingdom",
                 "Spain", "Italy", "Canada", "Australia")

@@

pisa_all2$value <- map(pisa_all2$value, ~ {
  .x$father_edu <- car::recode(as.numeric(.x$FISCED), "8:9 = NA")
  .x$mother_edu <- car::recode(as.numeric(.x$MISCED), "8:9 = NA")
  .x$high_edu_broad <- pmax(.x$father_edu, .x$mother_edu)
  .x$country <- pisa_countrynames[as.character(.x$CNT)]

  if (any(unique(.x$wave) %in% c("pisa2012", "pisa2015"))) {
    .x$father_edu <- .x$father_edu + 1
    .x$mother_edu <- .x$mother_edu + 1
    .x$high_edu_broad <- .x$high_edu_broad + 1
  }
  .x
})

reliability_pisa <-
  c("2000" = 0.81,
    "2003" = 0.85,
    "2006" = 0.78,
    "2009" = 0.74,
    "2012" = 0.82,
    "2015" = 0.74) # 2015 imputed

@@

  escs_trend <-
    map(escs_trend, ~ {
    mutate(.x, cnt = pisa_countrynames[cnt]) %>%
    rename(country = cnt)
  })

@@

# Function calculates the bottom 30th quantile for the bottom educated and the 70th quantile
# for the top educated. If the quantiles can't be estimated, it returns two NA's instead
quantile_missing <- function(df, weights) {

    df_lower <- filter(df, new_hisced %in% 1)
    df_upper <- filter(df, new_hisced == 7)

    quan_lower <- try(Hmisc::wtd.quantile(df_lower$escs_trend,
                                          weights = df_lower[[weights]],
                                          probs = c(0.50)))

    quan_upper <- try(Hmisc::wtd.quantile(df_upper$escs_trend,
                                    weights = df_upper[[weights]],
                                    probs = c(0.70)))

    if (any("try-error" %in% c(class(quan_lower), class(quan_upper)))) {
      return(c(NA, NA))
      } else {
     return(c(quan_lower[1], quan_upper[1]))
    }
}

@@

# It returns a dataframe for each survey with all countries and respective coefficients and
# standard errors.
test_diff <- function(df, reliability, test) {

@@

        # In some countries the quan can't be estimated because of very few obs.
        # The function doesn't stop but returns two NA's.
        quan <- quantile_missing(country, weights_var)

@@

      .x <-
        .x %>%
        select(wave, matches(paste0("^PV.*", test, "$")), escs_dummy,
               country, one_of(weights_var),
               AGE)

@@

      message(paste0(unique(.x$wave), " modeling done"))
      results
    })
}
@

@@

results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
# results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# US is missing for reading

# Cache is not working properly for the code above, so I just load the saved cached file
# load("./paper/cache/modeling_9a0b38d1d53fa243b0242580f0672fa5.RData")

<<>>=
vals <- c("Australia",
          "Germany",
          "Denmark",
          "Spain",
          "France",
          "Italy",
          "Netherlands",
          "Sweden",
          "United States")
# i <- 3
# vals <- intersect(unique(adapted_year_data[[1]]$country), unique(adapted_year_data[[2]]$country))

# while (i < 7) {
#   vals <- intersect(vals, unique(adapted_year_data[[i]]$country))
#   i <- i + 1
# }

sample_size_calc <- function(df, selected = F, cnts) {

@@

      split_df_two <-
        map(split_df, ~ {
      # In some countries the quan can't be estimated because of very few obs.
      # The function doesn't stop but returns two NA's.
      quan <- quantile_missing(.x, weights_var)

      # It's very important to create a variable that returns the number of observations of this dummy
      # For each country. Possibly to weight by the number of observations.
      .x$escs_dummy <-
        with(.x, case_when(escs_trend >= quan[2] ~ 1,
                                  escs_trend <= quan[1] ~ 0))
      .x
    })
  unsplit_df <- split_df_two %>% enframe() %>% unnest(value)
  unsplit_df
})

selected_cnt <- map(cnt_to_bind, ~ select(.x, country, high_edu_broad, new_hisced, escs_dummy))

selected_cnts
}

selected_cnts <- sample_size_calc(adapted_year_data, vals)  
xtable::xtable(selected_cnts)
@

<<>>=
all_cnts <- sample_size_calc(adapted_year_data)
@

>>>>

<<default_conf>>=
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # For pisa_countrynames
  library(countrycode) # For region variable
  library(car)
  library(readr)
  library(SAScii)
  library(inequalityintsvy)
  library(lme4)
  library(modelr)
  library(tidyverse)
  library(ggrepel)

@@

  countries <- c("Australia",
                 "Germany",
                 "Denmark",
                 "Spain",
                 "France",
                 "Italy",
                 "Netherlands",
                 "Sweden",
                 "Finland",
                 "United States",
                 "United Kingdom")
@

@@

<<loading_data-recoding, cache = TRUE>>=
pisa_all <- read_rds("./data/pisa_listcol.Rdata")
pisa_all2 <- pisa_all

years <- seq(2000, 2015, 3)

@@

pisa_all2$value <- map(pisa_all2$value, ~ {

# 2000 to 2015
# The coding is from 0 to 6, where 0 is no schooling and 6 is
# BA or above.

# When turning 0:6 to numeric, it becomes 1:7 that's why
# I recode 8:9 to NA. This, however, didn't work for last two surveys

  .x$father_edu <- car::recode(as.numeric(.x$FISCED), "8:9 = NA")
  .x$mother_edu <- car::recode(as.numeric(.x$MISCED), "8:9 = NA")
  .x$high_edu_broad <- pmax(.x$father_edu, .x$mother_edu)
  .x$country <- pisa_countrynames[as.character(.x$CNT)]

  if (any(unique(.x$wave) %in% c("pisa2012", "pisa2015"))) {
    # These two surveys were from 0:6 so I had to add + 1
    # so that it equals 1:7 as all other surveys.
    .x$father_edu <- .x$father_edu + 1
    .x$mother_edu <- .x$mother_edu + 1
    .x$high_edu_broad <- .x$high_edu_broad + 1
  }
  .x
})

reliability_pisa <-
  c("2000" = 0.81,
    "2003" = 0.85,
    "2006" = 0.78,
    "2009" = 0.74,
    "2012" = 0.82,
    "2015" = 0.74) # 2015 imputed
@

@@

  escs_trend <-
    map(escs_trend, ~ {
    mutate(.x, cnt = pisa_countrynames[cnt]) %>%
    rename(country = cnt)
  })
@

@@

# Function calculates the bottom 30th quantile for the bottom educated and the 70th quantile
# for the top educated. If the quantiles can't be estimated, it returns two NA's instead
quantile_missing <- function(df, weights, probs) {

    quan <- try(Hmisc::wtd.quantile(
      df$escs_trend,
      weights = df[[weights]],
      probs = probs
      ))

    if (any("try-error" %in% class(quan))) {
      return(c(NA, NA))
      } else {
     return(c(quan[1], quan[2]))
    }
}

@@

# It returns a dataframe for each survey with all countries and respective coefficients and
# standard errors.
test_diff <- function(df, reliability, test, probs) {

@@

        quan <- quantile_missing(country, weights_var, probs)

@@

      .x <-
        .x %>%
        dplyr::select(wave,
                      matches(paste0("^PV.*", test, "$")),
                      escs_dummy,
                      country,
                      one_of(weights_var),
                      AGE)

@@

      message(paste0(unique(.x$wave), " modeling done"))
      results
    })
}

# Adapted from: https://github.com/jtleek/slipper/blob/master/R/slipper.R
# Returns a tibble with the actual expr + the bootstrapped expr.
bootstrapper <- function(df, expr, B = 100, n = nrow(df), replacement = TRUE) {
  bootstrapper_(df, lazyeval::lazy(expr), B, n, replacement)
}

bootstrapper_ <- function(df, expr, B = 500, n = nrow(df), replacement = TRUE) {
  obs_val = lazyeval::lazy_eval(expr, data = df)
  boot_val = replicate(B, {
    newdata = sample_n(df, n, replace = replacement)
    lazyeval::lazy_eval(expr, data = newdata)
  })
  out = tibble(type = c("observed", "bootstrap"), 
               value = c(obs_val, mean(boot_val, na.rm = T)))
  return(out)
}

# For example
# bootstrapper(mtcars, mean(mpg), B = 200)
@

% This chunk needs to have cache = TRUE once you want to run it for like 2 hours until 
% the results_* models have been created once and the cache can save them

@@

# results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
# results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# results_math_topmid <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.5, 0.9))
# results_read_topmid <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.5, 0.9))
# results_math_midbottom <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.1, 0.5))
# results_read_midbottom <- test_diff(adapted_year_data, reliability_pisa, "READ", c(0.1, 0.5))

results_math <- read_rds("./data/delete.Rdata")
results_read <- read_rds("./data/delete_read.Rdata")
results_math_topmid <- read_rds("./data/delete_math_topmid.Rdata")
results_read_topmid <- read_rds("./data/delete_read_topmid.Rdata")
results_math_midbottom <- read_rds("./data/delete_math_midbottom.Rdata")
results_read_midbottom <- read_rds("./data/delete_read_midbottom.Rdata")
# US is missing for reading

# Cache is not working properly for the code above, so I just load the saved cached file
# load("./paper/cache/modeling_9a0b38d1d53fa243b0242580f0672fa5.RData")
@

% Figure out how to ouput data frames as latex tables.
<<sample_size>>=

# Get sample counts for each dummy
sample_size_calc <- function(df, probs, selected = F, cnts = NULL) {

  stopifnot(selected & !is.null(cnts))

@@

      split_df_two <-
        map(split_df, ~ {
          # In some countries the quan can't be estimated because of very few obs.
          # The function doesn't stop but returns two NA's.
          quan <- quantile_missing(.x, weights_var, probs)

          # It's very important to create a variable that returns the number of observations of this dummy
          # For each country. Possibly to weight by the number of observations.
          .x$escs_dummy <-
            with(.x, case_when(escs_trend >= quan[2] ~ 1,
                               escs_trend <= quan[1] ~ 0))
          .x
        })
      unsplit_df <- split_df_two %>% enframe() %>% unnest(value)

      unsplit_df %>%
        count(country, escs_dummy) %>%
        filter(!is.na(escs_dummy)) %>%
        left_join(summarize(group_by(unsplit_df, country), total_n = n()), by = "country") %>%
        mutate(perc = paste0(round(n / total_n * 100, 0), "%")) %>%
        select(-total_n)
    })
  setNames(cnt_to_bind, seq(2000, 2015, 3)) %>%
    enframe() %>%
    unnest()
}

sample_tables_topbottom <- sample_size_calc(adapted_year_data, c(.1, .9), selected = TRUE, countries)
sample_tables_topmid <- sample_size_calc(adapted_year_data, c(.5, .9), selected = TRUE, countries)
sample_tables_midbottom <- sample_size_calc(adapted_year_data, c(.1, .5), selected = TRUE, countries)
@
